A Computational Account of Default-Mode Function\\by Control Theory and Reinforcement Learning

### 2016-07-16 paper.tex
more TPJ anat notes

\begin{abstract}
ABC
%

% OUR keywords
%\textbf{\\keywords}: curse of dimensionality; semi-supervised learning;
%fMRI; systems neuroscience

@@

\section{Introduction}
%
\paragraph{The Human Default Mode Network.}
\begin{itemize}
  \item includes brain regions with highest baseline energy consumption in humans
  \item most active in non-disturbed mind-wandering without external influence
  \item tends to deactivate when humans engage in goal-direct tasks
  \item fluctuations in the DMN are related to lapses during and performance in
  such externally structured/focused tasks
  \item includes brain regions that come online first when waking up from
  anesthesia
  \item its network nodes are late to myelinate, an indicator of cognitive sophistication
  \item includes the most advanced processes hierarchies and has
  no direct connections with sensory regions processing external input
  \item the DMN's conceivable key role in the continuous environmental
  tracking in a generative, integrative process might explain both its highest energy consumption in the brain and its intimate coupling with conscious awareness
  \item Indeed, patients with right IPL (=DMN) damage have particular difficulties
  with multistep actions (Hartmann et al. 2005)
  and with imagined action (Sirigu et al. 1996).
  \item Lesion of the hippocampus, feeding memory and spatial information, to the
  DMN impairs future and hypothetical thinking (cf. Hassabis PNAS)
  \item a lesion study on disturbed sleep (i.e., a state of mind
  independent of sensory stimulation but dependent on
  internally generated information) exclusively identified
  the dmPFC (Koenigs et al., 2010). Third, another VLSM study exclusively
  related the IFG and TPJ, both more strongly connected to the
  dmPFC in our study, to inner speech
  \item Bálint's syndrome: neurological disorder of conscious
  awareness resulting from damage in bilateral parietal cortex;
  "psychic paralysis of gaze", that is, the inability for internally
  motivated, targeted saccades hindering fixation of currently not
  attended features in the field of view (i.e., oculomotor apraxia).
  unable to bind various individual features of the visual
  environment into an integrated whole (i.e., simultanagnosia)
  as well as unable to navigate hand movement to a targeted
  object helped by vision (i.e., optic ataxia).
  difficulty avoiding running into objects placed ahead;
  challenged in estimating distances as well as perceiving
  spatial depth, object sizes and object orientation
\end{itemize}

@@

@@

@@

relevance is the critical factor that determines whether an object activates the ventral network (Downar et al., 2001). The ventral network might be considered a prime candi-date for mediating orienting to salient but unimportant stimuli, i.e., exogenous attention (Posner and Cohen, 1984), be- cause under passive conditions it is highly responsive to distinctive sensory
events in all modalities (Downar et al., 2000). The ventral network is well activated by stimuli that are important, even if they are not very distinctive. 
CURRENT OPINION: VA network does underlie  stimulus-driven reorienting to environmental stimuli based on their task relevance. An important conclusion from these neuroimaging studies is that the psychological dis- tinction between exogenous and endogenous orienting (Jo- nides, 1981) may not map onto different neural systems. "changing the control of behavior from one environ- mental input to another". Similar reorienting mechanisms may also be involved in shifting from a broad range of ‘‘internally directed’’ processes in order to deal with environmental events, as when interrupting memory retrieval (‘‘did I lock the car door?’’) to respond to a sudden stimulus (‘‘is that my cell phone ring- ing?’’). -> medites between the Default-Mode function and other more environmentally directed processing goals.

\item LTPJ:

@@

\item
coding, transfer, and consolidation of
information
This emerging picture of hippocampal dy- namics suggests that neurons at the opposite ends of the distributions may convey differ- ent but complementary types of information. The ever-active minority of place cells may be responsible for generalizing across envi- ronments and affords the brain the capacity to regard no situation as completely un- known because every alley, mountain, river, or room has elements of previously experi- enced similar situations. In many situations,
this minority provides the “best guess” of the hippocampus and offers “good enough” solutions to get by. On the other hand, the majority of less active neurons constitute a large reservoir that can be mobilized (7) to precisely distinguish one situation from an- other and incorporate novel ones as distinct.
\end{itemize}
\end{itemize}

@@

\paragraph{How many controls do we need at best ?}
In the thermodynamic limit ($n \rightarrow \infty$), statistical physics \cite{Liu2011} gives the extremely good estimate
\begin{equation}
  n_0(\langle k\rangle, \gamma) \approx \exp\left(-\frac{1}{2}\left(\frac{\gamma-2}{\gamma - 1}\right)\langle k \rangle\right),
\end{equation}
where $\gamma \ge 1$ is the \textit{scale-free} parameter for the node-degree distribution of the network described by $\A$, and $\langle k \rangle$ is the mean degree. Letting $\gamma \rightarrow \infty$, one recovers the Erdos-Renyi value
\begin{equation}
  n_0(\langle k\rangle) = \exp\left(-\frac{1}{2}\langle k\rangle\right),
\end{equation}
which agrees perfectly with its known analytic value. One identifies two radically different regimes:
\begin{itemize}
  \item the manageable regime $\gamma > \gamma_c := 2$, in which $n_0 < 1$, and so we only need fewer controls than total number of nodes, and
\item the unmanageable regime ``$\gamma \le \gamma_c$'', in which  $n_0 = 1$, where each node must be controlled explicitly.
\end{itemize}

@@

\subsection{A thermodynamic model for bounded rationality, aka robust optimality}
\begin{itemize}
  \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
  \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
  \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
    \end{itemize}
The material presented is a revisit of \textcolor{blue}{[Braun 2011]}.
\paragraph{Utility functions and conjugate pairs.}
  Let $(\Omega, \mathcal F, P)$ be a probablity space. A function $U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $U(A|B) := U(A \cap B) - U(B)$ has the following propertites:
  \begin{itemize}
  \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.    
  \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
    $A,B \in \mathcal F$.
  \item monotonicity: $f_{U}$ is strictly increasing.
    \end{itemize}

\begin{theorem}
  The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $U(A|B) \equiv f(P(A|B))$ any probability space $(\Omega, \mathcal F, P)$ and utility function $U$ thereupon are of the form
  \begin{equation}
    f = \alpha \log(.),
    \label{eq:boltzmann}
  \end{equation}
  where $\alpha > 0$.
\end{theorem}

@@

Such $U$ and $P$ are said to form a \textit{conjugate pair} at temperature $\alpha$.

\paragraph{Example.}
Given a utility function $U$ on a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-U(\omega))_{\omega \in \Omega}$ is defined to be the probability measure (on thesame measurable space)
  \begin{equation}
    P(\omega) = \frac{1}{Z_{U}(\alpha)}\exp\left(\frac{1}{\alpha}U(\omega)\right), \; \forall \omega \in \Omega,
  \end{equation}
  where
  $Z_{U}(\alpha) := \sum_{\omega \in \Omega}\exp\left(\frac{1}{\alpha}U(\omega)\right)$ is a normalization constant called the \textit{partition function} of $U$.
  It's not hard to see that $U$ and the $P$ above form a conjugate pair.

  \paragraph{Free-utility functional.}
    Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$. Given another probability measure $P'$ on the same space, define it's \textit{free utility} as
    \begin{equation}
      J(P'|U, P) = \mathbb E_{P'}(U) + \alpha \mathbb H(P'),
    \label{eq:free_u}
    \end{equation}
    where $\mathbb H(P') := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))$ is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
  \begin{equation}
    J(P'|U,P) \le J(P|U,P) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
  \end{equation}
  In particular, if $P$ is the Gibbs measure corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
  \begin{equation}
    J(P'|U,P) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
    \end{equation}

  \paragraph{The free-energy / utility principle (of Friston ?).}
  We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it. Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$. We want to transform this to a new model by adding constraints represented by the utility function $\Delta U$.  The resulting system has final utility $U_{\text{fin}} = U_{\text{ini}} + \Delta U$. The difference in free-utility is then
  \begin{equation}
    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{P_{\text{fin}}}(\Delta U)}_{\textbf{accuracy}} + \alpha \underbrace{D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
    \label{eq:free}
  \end{equation}
  where $D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})$ is the Kullback-Leibler divergence.
  In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
  The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
a trade-off between an expected utility $\Delta U$ (first term) and
the information cost of transforming Pi
into $P_{\text{fin}}$ (second
term). In this interpretation $P_{\text{ini}}$ represents an initial choice
probability or policy, which includes the special case of the
uniform distribution where the decision-maker has initially no
preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
considers the utility constraint U∗
that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
\begin{equation}
P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
  \end{equation}

By differentiating the RHS of \eqref{eq:free} w.r.t $P_{\text{fin}}$ and setting to zero, we obtain the closed-form solution

\begin{equation}
  P^*_{\text{fin}}(\omega) \propto P_{\text{ini}}(\omega)\exp\left(\frac{1}{\alpha}\Delta U(\omega)\right).
\end{equation}

Two limit cases are worth considering.
\begin{itemize}
\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \mathbb E_{P'_{\text{fin}}}(\Delta U)$, and so it's optimal to take $P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*)$, where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
  \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
  \end{itemize}
\section{Experimental Results}
\paragraph{Serial versus parallel structure discovery and classification.}

@@

>>>>

\begin{abstract}

@@

network that might be involved in the automated prediction of social events that modulate behavior

\section{Introduction to the Default-Mode Network}
%
\paragraph{The Human Default Mode Network.}
scene construction processes when envisioning past, ficti- tious, and future events (Hassabis et al., 2007; Spreng et al., 2009).

It was observed that the more external stimuli are predictable, the more reflection processes become detached from the actual sensory environment and the more stimulus-independent, self-focused thoughts occur (Mason et al., 2007).

\begin{itemize}
  \item includes brain regions with highest baseline energy consumption in humans
  \item most active in non-disturbed mind-wandering without external influence
  \item tends to deactivate when humans engage in goal-direct tasks
  \item fluctuations in the DMN are related to lapses during and performance in
  such externally structured/focused tasks
  \item includes brain regions that come online first when waking up from
  anesthesia
  \item its network nodes are late to myelinate, an indicator of cognitive sophistication
  \item includes the most advanced processes hierarchies and has
  no direct connections with sensory regions processing external input
  \item the DMN's conceivable key role in the continuous environmental
  tracking in a generative, integrative process might explain both its highest energy consumption in the brain and its intimate coupling with conscious awareness
  \item Indeed, patients with right IPL (=DMN) damage have particular difficulties
  with multistep actions (Hartmann et al. 2005)
  and with imagined action (Sirigu et al. 1996).
  \item Lesion of the hippocampus, feeding memory and spatial information, to the
  DMN impairs future and hypothetical thinking (cf. Hassabis PNAS)

  \item
  autobiographical memory sup- plies numerous building blocks of social semantic knowl- edge (Bar 2007; Binder et al. 2009). These isolated conceptual scripts may be reassembled to enable the fore- casting of future events (Tulving 1983, 1985; Schacter et al. 2007). Similar brain mechanisms in remembering past episodes and envisioning future circumstances is supported by their engagement of identical brain areas, as evidenced by a quantitative meta-analysis (Spreng et al. 2009). Moreover, retrograde amnesic patients were repor- ted to be impaired not only in prospection but also in imagining novel experience (Hassabis et al. 2007). These findings suggest a single neural network for mentally constructing plausible semantic scenarios of detached sit- uations regardless of temporal orientation (Buckner and Carroll 2007; Hassabis and Maguire 2007). Indeed, con- struction of detached probabilistic scenes has been argued to influence ongoing decision making by estimating out- comes of behavioral choices (Boyer 2008; Suddendorf and Corballis 2007; Schilbach et al. 2008b). 
\end{itemize}

@@

PFC
Patients with prefrontal lesions have problems in adapting to novel situations and stimuli
Stuss, D. T. & Benson, D. F. The Frontal Lobes (Raven, New York, 1986).

@@

\item
\item Bálint's syndrome: neurological disorder of conscious
awareness resulting from damage in bilateral parietal cortex;
"psychic paralysis of gaze", that is, the inability for internally
motivated, targeted saccades hindering fixation of currently not
attended features in the field of view (i.e., oculomotor apraxia).
unable to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as unable to navigate hand movement to a targeted
object helped by vision (i.e., optic ataxia).
difficulty avoiding running into objects placed ahead;
challenged in estimating distances as well as perceiving
spatial depth, object sizes and object orientation

@@

relevance is the critical factor that determines whether an object activates the ventral network (Downar et al., 2001). The ventral network might be considered a prime candi-date for mediating orienting to salient but unimportant stimuli, i.e., exogenous attention (Posner and Cohen, 1984), be- cause under passive conditions it is highly responsive to distinctive sensory
events in all modalities (Downar et al., 2000). The ventral network is well activated by stimuli that are important, even if they are not very distinctive. 
CURRENT OPINION: VA network does underlie  stimulus-driven reorienting to environmental stimuli based on their task relevance. An important conclusion from these neuroimaging studies is that the psychological dis- tinction between exogenous and endogenous orienting (Jo- nides, 1981) may not map onto different neural systems. "changing the control of behavior from one environ- mental input to another". Similar reorienting mechanisms may also be involved in shifting from a broad range of ‘‘internally directed’’ processes in order to deal with environmental events, as when interrupting memory retrieval (‘‘did I lock the car door?’’) to respond to a sudden stimulus (‘‘is that my cell phone ring- ing?’’). -> medites between the Default-Mode function and other more environmentally directed processing goals.
inducing network reconfiguration

\item
blanke:
the core region of the human vestibular cortex is situated close to the angular gyrus10. It is possible that the experience of dissociation of self from the body is a result of failure to integrate complex somatosensory and vestibular information.
Initial stimulations (n 3; 2.0–3.0 mA) induced vestibular responses, in which the patient reported that she was “sinking into the bed” or “falling from a height”.

\item
high-level social cognition also engenerded expectation violations
-> that's why coincides with "agency" meta-analytic location
(Decety 2007):
suggest that activation in the TPJ during social cognition may therefore rely on a lower-level computational mech- anism involved in generating, testing, and correcting internal predictions about external sensory events.

\item
Sridharan: RTPJ involved in bottom-up saliency detection

\item
Corbetta2002:
 We suggest that the TPJ is also strongly engaged by stimuli that are behaviourally relevant but require a change in the current task set. 
 -> exogeneously induced state transition?
 <=> MCMC in HC / PCC -> slow? endogeneous changes in state transitions by the policy and value matrices
 For example, the alarm at the museum is not part of the task set of listening to the guide’s discussion of Hieronymous Bosch, but it is clearly a behaviourally relevant stimulus.

\item
visuo- proprioceptive conflict (Balslev et al., 2005), and multi-modal detection of sensory changes (Downar et al., 2000)
brain areas connected to aRTPJ further responded to all changes of visual, auditory, or tactile stimulation in a multi-modal fMRI study (Downar et al., 2000).

\item
direct electrical stimulation of the RTPJ during neurosurgery was associated with altered perception and stimulus awareness (Blanke et al., 2002).

\item
the aRTPJ network is central for sensorimotor control by integrating supramodal stimulus-guided attention and action initia- tion during externally structured tasks.

\item
Bzdok2013

pRTPJ cluster was also functionally associated with deception tasks

qualify RTPJ as a po- tential switch between exteroceptive and interoceptive mind sets implemented by their networks. The reciprocal nature of those mind sets is well illustrated by consistent reports of continuous shifting be- tween externally (i.e., sensory) and internally (i.e., self- and presumably social/memory-) oriented processing in controlled laboratory settings and daily routine (Smallwood et al., 2007).

\item
the difference between the two task groups could be parsimoniously explained by the required atten- tion to either the external world or self.
world -> policy matrix driven computation
self -> value matrix driven computation

\item
significantly reduced impact of intentions after transient TPJ disruption (Young et al. 2010b)

LTPJ:

\item
only the left TPJ was associated with language processes, whereas only the right TPJ was functionally associated with attention processes

\item 
Wernicke's area:
involved in accessing and integrating world knowlege.
involved in the comprehension or understanding of written and spoken language Damage caused to Wernicke's area results in receptive, fluent aphasia. This means that the person with aphasia will be able to fluently connect words, but the phrases will lack meaning.
A major impairment of language comprehension, while speech retains a natural-sounding rhythm and a relatively normal syntax. Language as a result is largely meaningless. difficulty finding words, putting words in the wrong order, substituting sounds and/or words, difficulty understanding spoken utterances, difficulty recognizing some words by sight, difficulty writting letters

\item a lesion study on disturbed sleep (i.e., a state of mind
independent of sensory stimulation but dependent on
internally generated information) exclusively identified
the dmPFC (Koenigs et al., 2010). Third, another VLSM study exclusively
related the IFG and TPJ, both more strongly connected to the
dmPFC in our study, to inner speech

\item
the left and right TPJ further differ according to their cytoarchitectonic borders and gyral pattern (Caspers et al., 2008; Caspers et al., 2006)

@@

\item
coding, transfer, and consolidation of
information
This emerging picture of hippocampal dy- namics suggests that neurons at the opposite ends of the distributions may convey differ- ent but complementary types of information. The ever-active minority of place cells may be responsible for generalizing across envi- ronments and affords the brain the capacity to regard no situation as completely un- known because every alley, mountain, river, or room has elements of previously experi- enced similar situations. In many situations,
this minority provides the “best guess” of the hippocampus and offers “good enough” solutions to get by. On the other hand, the majority of less active neurons constitute a large reservoir that can be mobilized (7) to precisely distinguish one situation from an- other and incorporate novel ones as distinct.

\item
damage to even the hippocampus alone is sufficient to cause amnesia +
envisoning themselves in future and even hypothetical scenarios/mentally constructing new experiences

The HC, in turn, is well known to be involved in memory and spatial navigation in animals and humans (von Bechterew, 1900; Scoville and Milner, 1957; O’Keefe and Dostrovsky, 1971; Maguire et al., 2000).

\item
Hassabias2007:
Recollection of these rich autobiographical or episodic memories has been likened to mentally traveling back in time and re-experiencing one’s past (1). It has long been known that the hippocampus and related medial temporal lobe structures play a critical role in supporting episodic memory

In fact, episodic memory and imagining or constructing events share striking similarities in terms of the psychological processes engaged (19–21). These include imagery (22), sense of presence (1), retrieval of semantic information and multimodal details (23), and narrative structure (22).

visual imagery itself is not believed to be realized in the HC

@@

\paragraph{How many controls do we need at best ?}
In the thermodynamic limit ($n \rightarrow \infty$), statistical physics \cite{Liu2011} gives the extremely good estimate
\begin{equation}
  n_0(\langle k\rangle, \gamma) \approx e^{-\frac{1}{2}\left(\frac{\gamma-2}{\gamma - 1}\right)\langle k \rangle},
\end{equation}
where $\gamma \ge 1$ is the \textit{scale-free} parameter for the node-degree distribution of the network described by $\A$, and $\langle k \rangle$ is the mean degree. Letting $\gamma \rightarrow \infty$, one recovers the Erdos-Renyi value
\begin{equation}
  n_0(\langle k\rangle) = e^{-\frac{1}{2}\langle k\rangle},
\end{equation}
which agrees perfectly with its known analytic value. One identifies two radically different regimes:
\begin{itemize}
  \item the manageable regime $\gamma > \gamma_c := 2$, in which $n_0 < 1$, and so we only need fewer controls than total number of nodes, and
\item the unmanageable regime ``$\gamma \le \gamma_c$'', in which  $n_0 = 1$, where each node must be controlled explicitly.
\end{itemize}

@@

\subsection{A thermodynamic model for bounded rationality, aka robust optimality}
\begin{itemize}
  \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
  \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
  \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
    \end{itemize}

\paragraph{Utility functions and conjugate pairs.}
  Let $(\Omega, \mathcal F, \P)$ be a probablity space. A function $\U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $\U(A|B) := \U(A \cap B) - \U(B)$ has the following propertites:
  \begin{itemize}
  \item additivity: $\U(A_1 \cap A_2 | B) = \U(A_1|B) + \U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.    
  \item statistic: there exists a function $f_{\U} :\mathbb R_+ \rightarrow \mathbb R$ such that $\U(A|B) = f(\P(A|B))$, for all events
    $A,B \in \mathcal F$.
  \item monotonicity: $f_{\U}$ is strictly increasing.
    \end{itemize}

\begin{theorem}
  The only functions $f: \mathbb R_+ \rightarrow \mathbb R$ which is such that $\U(A|B) \equiv f(\P(A|B))$ any probability space $(\Omega, \mathcal F, \P)$ and utility function $\U$ thereupon are of the form
  \begin{equation}
    f = \alpha \log(.),
    \label{eq:boltzmann}
  \end{equation}
  where $\alpha > 0$.
\end{theorem}

@@

Such $\U$ and $\P$ are said to form a \textit{conjugate pair} at temperator $\alpha$.

\paragraph{Example.}
Given utility function $U$ a probability space $(\Omega, \mathcal F, *)$, the \textit{Gibbs measure} at temperature $\alpha > 0$ and energy levels $(-\U(\omega))_{\omega \in \Omega}$ is defined to be probability measure (on thesame measurable space)
  \begin{equation}
    \P_{\U}(A) = \sum_{\omega \in A}\frac{1}{Z(\alpha)}\exp\left(\frac{1}{\alpha}\U(\omega)\right), \; \forall A \in \mathcal F.
  \end{equation}
$Z(\alpha)$ is a normalization constant such that $\P_{\U}(\Omega) = 1$.  It's not hard to see that $\U$ and $\P_{\U}$  form a conjugate pair.

  \paragraph{Free-utility functional.}
    Let $(\U, \P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$. Given another probability measure $\P'$ on the same space, define it's \textit{free utility} as
    \begin{equation}
      J(\P'|\U, \P) = \mathbb E_{\P'}(\U) + \alpha \mathbb H(\P'),
    \label{eq:free_u}
    \end{equation}
    where $\mathbb H(\P') := -\sum_{\omega \in \Omega}\P'(\omega)\log(\P'(\omega))$ is the entropy of $\P'$.

  It's not difficult to show that
  \begin{equation}
    J(\P'|\U,\P) \le J(\P|\U,\P) = \sum_{\omega \in \Omega}\U(\omega) =: \U(\Omega).
  \end{equation}
  In particular, if $\P = \P_{\U}$ the Gibbs measure, then the upper bound above reduces to the \textit{log-partition function}
  \begin{equation}
    J(\P'|\U,\P) \le \U(\Omega) = -\alpha \log\left(\sum_{\omega \in \Omega}e^{\frac{1}{\alpha}\U(\omega)}\right).
    \end{equation}

  \paragraph{The free-energy / utility principle.}
  We are now in shape to introduce the notion of free-energy for transition of priors, and a variational principle for optimizing it. Consider thus an initial system described by a conjugate pair $(\U_{\text{ini}}, \P_{\text{ini}}) \rightarrow \P_{\text{fin}}$ at temperature $\alpha > 0$. We want to transform this to a new one by adding constraints represented by the utility function $\delta \U$.  The resulting system has final utility $\U_{\text{fin}} = \U_{\text{ini}} + \delta\U$. The difference in free-utility is then
  \begin{equation}
    \delta J = J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{\P_{\text{fin}}}(\delta \U)}_{\textbf{expected gain in utility}} + \alpha \underbrace{D_{\text{KL}}(\P_{\text{ini}}\|\P_{\text{fin}})}_{\textbf{complexity of change}},
    \label{eq:free}
  \end{equation}
  where $D_{\text{KL}}(\P_{\text{ini}}\|\P_{\text{fin}})$ is the Kullback-Leibler divergence.
  In the above formula, we've extensively used the fact that $(\U_{\text{ini}}, \P_{\text{ini}})$ is a congugate pair and so $\U_{\text{ini}}(\omega) \equiv \alpha \log(\P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
  The two terms in \eqref{eq:free} (expected gain in utility or negative workdone, and the complexity of the transition)

\section{Experimental Results}
\paragraph{Serial versus parallel structure discovery and classification.}

@@

formalize and predict
